---
title: "Egg CVD study"
output: github_document
always_allow_html: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "readxl", "lubridate", "tableone", "gridExtra", "survival")
sapply(pacs, require, character.only = TRUE)

# Function to search variables
search_var <- function(df, pattern, ...) {
  loc <- grep(pattern, names(df), ...)
  if (length(loc) == 0) warning("There are no such variables")
  else return(data.frame(loc = loc, varname = names(df)[loc]))
}

# Medicare crosswalk ------------------------------------------------------

# Read crosswalk file: n = 70.968
crosswalk <- read_fwf("./Data/12172/2022/ssn_bene_xwalk_res000058038_req012172_2022.dat",
                      fwf_widths(c(9, 15, 1, 1, 1), c("ORIG_SSN", "BENE_ID", "SSN_MATCH", "SEX_MATCH", "DOB_MATCH")))

# Extract matched BENE_IDs
all_matched_bene_ids <- crosswalk %>% 
  filter(SSN_MATCH == 1)

# BENE_IDs of gender or DOB mismatch (to be removed)
mismatches <- crosswalk %>% 
  filter(SSN_MATCH == 1 & (SEX_MATCH == 0 | DOB_MATCH == 0)) %>%
  select(BENE_ID)
  
# There are two duplicates in BENE_IDs...
dup_BENE_IDs <- all_matched_bene_ids %>% 
  group_by(BENE_ID) %>% 
  summarize(n = sum(n())) %>% 
  filter(n > 1)

# There are 106 SSN duplicates 
dup_SSNs <- all_matched_bene_ids %>% 
  group_by(ORIG_SSN) %>% 
  summarize(n = sum(n())) %>% 
  filter(n > 1)

# BENE_IDs that need to be removed:
# Gender/DOB mismatch
# SSN/BENE_ID duplicates
exclude_BENE_IDs <- dup_BENE_IDs %>% 
  select(BENE_ID) %>% 
  union(all_matched_bene_ids %>% 
          filter(ORIG_SSN %in% dup_SSNs$ORIG_SSN) %>% 
          select(BENE_ID)
  ) %>% 
  union(mismatches)

# Medicare MSBF file ------------------------------------------------------

# Read MBSF Summary data on each year
# Data specification of MSBF files
fts_msbf <- read_excel("./Data/mbsf_format.xlsx")

# Create file names
year <- 2008:2020
fname <- paste0("./Data/12172/", year, "/mbsf_abcd_summary_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_msbf <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(fts_msbf$length, fts_msbf$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)

# Long format over years
all_msbf_long <- all_msbf %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, BENE_ENROLLMT_REF_YR)

# Medicare chronic condition file -----------------------------------------

# Read chronic conditions data on each year
fts_cc <- read_excel("./Data/mbsf_cc_format.xlsx")

# Create file names
year <- 2008:2020
fname <- paste0("./Data/12172/", year, "/mbsf_cc_summary_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_cc <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(fts_cc$length, fts_cc$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)

# Long format over years
all_cc_long <- all_cc %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, BENE_ENROLLMT_REF_YR)

# AHS-2 Medicare link file ------------------------------------------------

# AHS-2 analysis ID: N = 51,917
ahs <- read_csv("./Data/MedicareMatches2022.csv")

# There are 103 analysis IDs that appears twice
# There are 226 NULL values on analysis ID
# 104 analysis IDs (103 duplicates + NAs) to be removed
exclude_analysisIDs <- ahs %>% 
  group_by(AnalysisID) %>%
  tally() %>% 
  filter(n > 1) %>%
  select(AnalysisID)

# 432 rows removed (103 * 2 + 226)
# yielding 51,485 unique analysis ID (and unique Bene_IDs)
ahs_dup_removed <- ahs %>% anti_join(exclude_analysisIDs)

# AHS-2 data --------------------------------------------------------------

# N = 96,144
ahsdata <- read.csv("./Data/BaselineDataForMedicare20190531.csv", header=TRUE)
names(ahsdata) <- tolower(names(ahsdata))

# Remove those with missing analysis ID
# Calculate birth date and age in 2010
ahsdata <- 
  ahsdata %>% 
  filter(!is.na(analysisid)) %>% 
  mutate(mob         = ifelse(mob > 12, NA, mob),
         dob         = ifelse(dob > 31, NA, dob),
         yob         = ifelse(yob >= 9999, NA, yob),
         bdate       = as.Date(paste(yob, mob, dob, sep="-")),
         qreturndate = as.Date(qreturndate),
         agein       = as.numeric((qreturndate - bdate) / 365.242),
         age2010     = floor(as.numeric((as.Date("2010-12-31") - bdate) / 365.242)),
         age2011     = floor(as.numeric((as.Date("2011-12-31") - bdate) / 365.242)),
         deceased    = as.Date(deceased, format = "%Y-%m-%d"))

# Convert to numeric
ahsdata <- ahsdata %>% 
  mutate_at(vars(eggbetrf:educyou, incomeh:alcnow, cancer, anginay:a05q9, sleephrs:walkruna), 
            function(x) as.numeric(as.character(x)))

# Calculate BMI and BMI category
ahsdata <- ahsdata %>% 
  mutate(heighti = ifelse(heighti > 11, NA, heighti),
         bmi     = weight / (heightf * 12 + heighti) ^ 2 * 702,
         # bmicat  = cut(bmi, breaks = c(0, 18.5, 25, 30, Inf), right = FALSE),
         bmicat  = cut(bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
         # bmicat  = factor(bmicat, labels = c("Underweight", "Normal", "Overweight", "Obese")))
         bmicat  = factor(bmicat, labels = c("Normal", "Overweight", "Obese")))

# Recode marital statas, education and sleep
ahsdata <- ahsdata %>% 
  mutate(marital   = recode(marital, "Never", "Married", "Married", "Married", "Div/Wid", "Div/Wid", "Div/Wid"),
         marital   = factor(marital, levels = c("Married", "Never", "Div/Wid")),
         educyou   = recode(educyou, "HS or less",   "HS or less",   "HS or less", 
                                     "Some college", "Some college", "Some college", 
                                     "Col grad+",    "Col grad+",    "Col grad+"),
         educyou   = factor(educyou, levels = c("HS or less", "Some college", "Col grad+")),
         educyou2  = relevel(educyou, ref = "Col grad+"),
         sleephrs  = recode(sleephrs, "<= 5 hrs", "<= 5 hrs", "<= 5 hrs", "6 hrs", "7 hrs",  
                                      "8 hrs", ">= 9 hrs", ">= 9 hrs", ">= 9 hrs"),
         sleephrs  = factor(sleephrs, levels = c("<= 5 hrs", "6 hrs", "7 hrs", "8 hrs", ">= 9 hrs")),
         sleephrs2 = relevel(sleephrs, ref = "7 hrs"))

# Dietary pattern
ahsdata$vegstat <- factor(ahsdata$vegstat, labels=c("Vegan", "Lacto-ovo",  "Semi", "Pesco", "Non-veg"))

# Physical activity
test <- ahsdata %>% 
  mutate(times    = walkrunf - 1,
         duration = recode(exeramt, 0, 5, 15, 25, 35, 45, 55, 60),
         minutes  = recode(walkruna, 5, 15, 25, 35, 45, 55, 60),
         distance = recode(walkrund, 0.25, 0.5, 1, 1.5, 2, 3, 4),
         vigofreq = ifelse(exerfreq - 2 < 0, 0, exerfreq - 2),
         ex4b4c   = times * distance,
         ex4b4d   = times * minutes,
         ex3a3b   = vigofreq * duration)

test <- test %>% 
  mutate(none = ifelse((ex3a3b == 0 & (walkrun == 2 | is.na(walkrun))) | 
                       (ex3a3b == 0 & times == 0) | (vigofreq == 0 & walkrun == 2), 1, 0),
         low  = ifelse((0 < ex3a3b & ex3a3b < 105) | (0 < ex4b4d & ex4b4d < 105) | (0 < ex4b4c & ex4b4c < 3), 1, 0),
         mid  = ifelse(ex3a3b >= 105 | (105 <= ex4b4d & ex4b4d < 175) | (3 <= ex4b4c & ex4b4c < 9), 1, 0),
         hi   = ifelse(ex4b4d >= 175 | ex4b4c >= 9, 1, 0))

ahsdata$exercise[test$none == 1] <- 1
ahsdata$exercise[test$low  == 1] <- 2
ahsdata$exercise[test$mid  == 1] <- 3
ahsdata$exercise[test$hi   == 1] <- 4
ahsdata$exercise <- factor(ahsdata$exercise, labels=c("None", "Low", "Moderate", "Vigorous"))
rm(test)

# AHS-2 dietary data ------------------------------------------------------

# N = 91099 obs, 105 variables
ahsdiet0 <- read.csv("./Data/MEDC_DMENT-FULL-20191008-0.csv", header = TRUE)

# N = 4498 obs, 105 variables
ahsdiet_ex <- read.csv("./Data/MEDC_DMENT-FULL-EXCLUDED-20191008-0.csv", header = TRUE)

# Concatenate two dfs
ahsdiet <- ahsdiet0 %>% bind_rows(ahsdiet_ex)
names(ahsdiet) <- tolower(names(ahsdiet))
ahsdiet$bmi <- NULL

# Remove those with "cannot use diet"
# Yields N = 95,344
ahsdiet <- ahsdiet %>% 
  filter(x_cannotusediet == 0)

# Merge AHS data with Medicare --------------------------------------------

# Extract data of the last seen: MSBF and CC files  
msbf_last_seen <- all_msbf_long %>%
  group_by(BENE_ID) %>% 
  slice(n())

cc_last_seen <- all_cc_long %>%
  group_by(BENE_ID) %>% 
  slice(n())

# Merge AHS data with Medicare
# Results in n = 44,159 subjects
ahs_medic <- msbf_last_seen %>% 
  inner_join(cc_last_seen %>% select(-BENE_ENROLLMT_REF_YR), by = "BENE_ID") %>%
  inner_join(ahs_dup_removed %>% 
               rename(BENE_ID = Bene_ID) %>% 
               mutate(analysisid = parse_number(AnalysisID)), by = "BENE_ID") %>% 
  inner_join(ahsdata, by = "analysisid") %>% 
  # left_join(ahsdiet, by = "analysisid") %>% 
  inner_join(ahsdiet, by = "analysisid") %>% 
  ungroup()

# Opt-outs: n = 395
optout <- read_csv("./Data/OptOutAnalysisIDs.csv") %>% 
  setNames("analysisid")

# Remove opt-outs, yielding n = 44,013
ahs_medic <- ahs_medic %>% 
  anti_join(optout, by = "analysisid")


# Apply inclusion/exclusion criteria --------------------------------------

# Remove if AGE_AT_END_REF_YR < 65 (n = 1537)
# Results in n = 42,476
ahs_medic <- ahs_medic %>% 
  filter(AGE_AT_END_REF_YR >= 65)

# Remove if BMI is missing (n = 1151) or extreme (n = 146)
# Resuts in n = 41,174
ahs_medic <- ahs_medic %>% 
  filter(bmi >= 16, bmi <= 60)

# Exclude extreme kcal intake (n = 1030)
# Results in n = 40,149
ahs_medic <- ahs_medic %>% 
  filter(kcal > 500 & kcal < 4500)

# Need to exclude unverified deaths
# There are 22 unverified deaths
unverified_deaths <- ahs_medic %>% 
  filter(!is.na(BENE_DEATH_DT)) %>%
  filter(is.na(VALID_DEATH_DT_SW)) %>% 
  select(analysisid)

# Exclude unverified deaths
# Yields n = 40,127
ahs_medic <- ahs_medic %>% 
  anti_join(unverified_deaths, by = "analysisid") 

```

## Aim

* Assess if the risk of developing CVD varies according to levels of meat intake among regular egg consumers with hyperlipidemia using the AHS-2 cohort linked with Medicare data

* Assess the interaction between race and egg intake on the incidence of CVD among those with hyperlipidemia. 

## Datasets

* Medicare data
  * For details regarding Medicare data, see [AHS-2 Medicare Linkage](https://github.com/keijioda/ahs_medicare_linkage/blob/main/summary.md) repository.
  
  * Master Beneficiary Summary File (MBSF), 2008-2020
    * Contains beneficiary characteristics and enrollment information
    
  * Chronic Conditions file (CC), 2008-2020
    * Contains the first occurrence date of 27 specific chronic conditions
    * Used to identify prevalent/incident cases of dementia and/or Alzheimer's disease and
    * to identify comorbidities
    
  * Both files include n = 44,585 unique subjects across years, after excluding 
    * Gender/DOB mismatch with AHS-2 data
    * Dupulicate beneficiary IDs and SSNs

* AHS-2
  * Baseline data: n = 96,144
  * Dietary data: n = 95,344
  
* After merging Medicare and AHS-2 data, there were n = 44,159 subjects.

* Among n = 44,159 subjects, there were n = 146 subjects who opted out from the study. Excluding these subjects yielded n = 44,013.

## Inclusion/exclusion criteria

* Medicare beneficiaries who did not reach the age of 65 between 2008 and 2020 (e.g., younger beneficiaries with disabilities or end-stage renal disease) were excluded (n = 1537), resulting n = 42,476.

* n = 1302 subjects with missing BMI or extreme BMI (<16 or >60), according to AHS questionnaire, were excluded, resulting n = 41,174.

* n = 1030 subjects with extreme calorie intake (<500 or >4500 kcal) were excluded, resulting n = 40,149.

* Unverified dates of deaths
  * Medicare data include a variable (`VALID_DEATH_DT_SW`) indicating whether a beneficiary's day of death has been verified by the Social Security Administration or the Railroad Retirement Board.
  * There were 22 unverified death dates. Excluding these resulted n = 40,127.

## Exploratory analysis

```{r hyperlipidemia, include = FALSE}
# Identify those with hyperlipidemia
# Both prevalence and incidence
ahs_medic <- ahs_medic %>% 
  mutate(HYPERL_YN = ifelse(is.na(HYPERL_EVER), 0, 1),
         HYPERL_YN = factor(HYPERL_YN, label = c("No", "Yes")),  
         BENE_BIRTH_DT = ymd(BENE_BIRTH_DT),
         BENE_DEATH_DT = ymd(BENE_DEATH_DT),
         HYPERL_EVER = ymd(HYPERL_EVER))

# There are 21,530 hyperlipidemia cases (53.7%)
ahs_medic %>%
  group_by(HYPERL_YN) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)

# Find prevalent cases
ahs_medic %>% 
  filter(HYPERL_YN == "Yes") %>% 
  mutate(HYPERL_EVER = ymd(HYPERL_EVER)) %>%
  filter(HYPERL_EVER < qreturndate) %>% 
  select(analysisid, HYPERL_YN, qreturndate, HYPERL_EVER)

hyperl_diag_date <- ahs_medic %>% 
  filter(HYPERL_YN == "Yes") %>% 
  mutate(HYPERL_EVER = ymd(HYPERL_EVER)) %>%
  mutate(DateDiff = interval(qreturndate, HYPERL_EVER),
         DateDiff_days = as.numeric(DateDiff, 'days'), 
         DateDiff_months = as.numeric(DateDiff, 'months'), 
         DateDiff_years = as.numeric(DateDiff, 'years')) 

# How many cases in which qreturndate > diagnosis date?
# 6,464 such cases
hyperl_diag_date %>% 
  filter(DateDiff_days < 0) %>% 
  select(HYPERL_YN, HYPERL_EVER, qreturndate, DateDiff_days, DateDiff_months, DateDiff_years)

# How many cases who were diagnosed within 6 months after qreturndate?
# 644 cases
hyperl_diag_date %>% 
  filter(between(DateDiff_months, 0, 6)) %>% 
  select(HYPERL_YN, qreturndate, HYPERL_EVER, DateDiff_days, DateDiff_months, DateDiff_years)

# Considering 6 months as a cut-off, there are 7,138 prevalent cases
hyperl_diag_date %>% 
  filter(DateDiff_months <= 6) %>% 
  select(BENE_ID, analysisid, HYPERL_YN, qreturndate, HYPERL_EVER, DateDiff_days, DateDiff_months, DateDiff_years)
```

### Hyperlipidemia cases
* There are 21,530 (53.7%) subjects who have developed hyperlipidemia before the end of year 2020
* Among them, there are 6,464 subjects whose diagnosis date precedes the enrollment into AHS-2
* Plus, there are 644 subjects who were diagnosed within 6 months after the enrollment
* Thus, considering 6 months as a cut-off, there are 7,108 prevalent cases (33.0% among all hyperlipidemia cases)
* [To consider]
    * We could use the status of hyperlipidemia (yes/no) as time-dependent variable in the cox model
    * Or include into analysis anyone with the diagnoses regardless of prevalence/incidence? 

### CVD cases
* The medicare data include five relevant diagnosis: Ischemic HD, stroke/TIA, acute MI, CHF, and atrial fibrillation
* For each of these 5 conditions, the number of cases (prevalence and incidence) is shown below:
  
```{r five_cvd_cc, echo = FALSE}
cvd_vars <- c("ISCHEMICHEART_EVER", "STROKE_TIA_EVER", "AMI_EVER", "CHF_EVER", "ATRIAL_FIB_EVER")

cvd_df <- ahs_medic %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), "No", "Yes"))

CreateTableOne(cvd_vars, data = cvd_df) %>% kableone()
```

* Many of these CVD cases have more than one of the 5 conditions. See below for the frequency table of CVD conditions:

```{r n_cvd_cc, echo = FALSE}
# Number of CVDs per subject
ahs_medic %>% 
  select(all_of(cvd_vars)) %>% 
  mutate_all(ymd) %>% 
  mutate_all(\(x) ifelse(is.na(x), 0, 1)) %>% 
  mutate(n_CVDs = rowSums(across(all_of(cvd_vars)))) %>% 
  group_by(n_CVDs) %>%
  tally() %>%
  mutate(pct = n / sum(n)) %>%
  knitr::kable()

# Based on ischemic HD, stroke, MI, CHF and Afib
ahs_medic2 <- ahs_medic %>% 
  # mutate(CVD_EVER = pmin(ISCHEMICHEART_EVER, STROKE_TIA_EVER, AMI_EVER, na.rm = TRUE),
  mutate(CVD_EVER = pmin(ISCHEMICHEART_EVER, STROKE_TIA_EVER, AMI_EVER, CHF_EVER, ATRIAL_FIB_EVER, na.rm = TRUE),
         CVD_EVER = ymd(CVD_EVER), 
         CVD_YN = ifelse(is.na(CVD_EVER), 0, 1),
         CVD_YN = factor(CVD_YN, labels = c("No", "Yes")))

# Find prevalent cases
cvd_diag_date <- ahs_medic2 %>% 
  filter(CVD_YN == "Yes") %>% 
  mutate(DateDiff = interval(qreturndate, CVD_EVER),
         DateDiff_days = as.numeric(DateDiff, 'days'), 
         DateDiff_months = as.numeric(DateDiff, 'months'), 
         DateDiff_years = as.numeric(DateDiff, 'years')) 

# How many cases in which qreturndate > diagnosis date?
# 5,297 such cases
prev_cvd <- cvd_diag_date %>% 
  filter(DateDiff_days <= 0) %>% 
  select(analysisid)
```

* If we define CVD cases as any one of the 5 conditions:
  * There are 18,105 cases (45.1%) of CVDs
  * Among them, there are 5,297 prevalent cases and 12,808 incident cases
  * When the prevalent CVD cases were excluded, we have an analytic sample of n = 34,830 with 12,808 incident CVD cases (36.8%).

* [Approach] Start with all 5 conditions combined and then explore individual conditions as sensitivity analysis

### Other variables

```{r other_vars, echo = FALSE}
# Define variables for models ---------------------------------------------

# Define ageout
# If CVD_EVER exists (incident cases), the use this diag date
# If non-case and BENE_DEATH_DT exists, then use this date died (censored)
# Otherwise, use the end of BENE_ENROLLMT_REF_YR (year last seen)
ahs_medic_inc <- ahs_medic2 %>% 
  anti_join(prev_cvd, by = "analysisid") %>% 
  mutate(
    age_last_seen = time_length(interval(BENE_BIRTH_DT, make_date(BENE_ENROLLMT_REF_YR, 12, 31)), "year"),
    ageout = case_when(
              CVD_YN == "Yes" ~ time_length(interval(BENE_BIRTH_DT, CVD_EVER), "year"),
              CVD_YN == "No" & !is.na(BENE_DEATH_DT)  ~ time_length(interval(BENE_BIRTH_DT, BENE_DEATH_DT), "year"),
              CVD_YN == "No" &  is.na(BENE_DEATH_DT)  ~ age_last_seen),
    fuyear = ageout - agein
  )

# Factor gender, categorize age into age groups, recode RTI race, smoking, and drinking
ahs_medic_inc2 <- ahs_medic_inc %>% 
  mutate(
    bene_sex_F = factor(SEX_IDENT_CD, labels = c("M", "F")),
    # bene_age_at_end_2008 = time_length(interval(BENE_BIRTH_DT, make_date(2008, 12, 31)), "year"),
    bene_age_at_end_2020 = time_length(interval(BENE_BIRTH_DT, make_date(2020, 12, 31)), "year"),
    # agecat     = cut(bene_age_at_end_2008, breaks = c(50, 60, 65, 70, 75, 80, 85, 120), right = FALSE),
    # agecat     = factor(agecat, labels = c("<60", "60-64", "65-69", "70-74", "75-79", "80-84", "85+")),
    agecat     = cut(bene_age_at_end_2020, breaks = c(65, 70, 75, 80, 85, 90, 95, 130), right = FALSE),
    agecat     = factor(agecat, labels = c("65-69", "70-74", "75-79", "80-84", "85-89", "90-94", "95+")),
    rti_race3  = recode(RTI_RACE_CD + 1, 3, 1, 2, 3, 3, 3, 3),
    rti_race3  = factor(rti_race3, labels = c("NH White", "Black", "Other")),
    smokenow   = ifelse(is.na(smokenow), 0, smokenow),
    smokecat   = ifelse(smokenow == 1, 2, ifelse(smoke > 1, 2, 1)),
    smokecat   = factor(smokecat, labels = c("Never", "Ever")),
    alcnow     = ifelse(is.na(alcnow), 0, alcnow),
    alccat     = ifelse(alcnow == 1, 2, ifelse(alcohol > 1, 2, 1)),
    alccat     = factor(alccat, labels = c("Never", "Ever")),
    vegstat2   = relevel(vegstat, ref = "Non-veg"))

# Co-morbidity ------------------------------------------------------------

# Prevalent comorbidity according to CMS
dzvars <- c("cataract", "chronickidney", "copd", "diabetes", "glaucoma",
            "hip_fracture", "depression", "osteoporosis", "ra_oa",  
            "cancer_breast", "cancer_colorectal", "cancer_prostate", "cancer_lung", "cancer_endometrial",
            "asthma", "hyperl", "hypert", "hypoth", "anemia")
dzvars <- paste0(toupper(dzvars), "_EVER")

# Convert all comodidity ever variables to date
test <- ahs_medic_inc2 %>% 
  mutate(across(all_of(dzvars), ymd)) 

find_prevalence <- function(var, data, start){
  cc_var <- data[[var]]
  strtdt  <- data[[start]]
  out <- ifelse(cc_var <= strtdt, 1, 0)
  out <- ifelse(is.na(cc_var), 0, out)
  return(out)
}

dzdf <- as.data.frame(lapply(dzvars, find_prevalence, data = test, start = "qreturndate"))
names(dzdf) <- paste0(dzvars, "_YN")
dzdf <- as_tibble(dzdf)

dzdf <- dzdf %>% 
  mutate(como_depress  = DEPRESSION_EVER_YN,
         como_diabetes = DIABETES_EVER_YN,
         como_kidney   = CHRONICKIDNEY_EVER_YN,
         como_hypoth   = HYPOTH_EVER_YN,
         como_anemia   = ANEMIA_EVER_YN,
         como_hypert   = HYPERT_EVER_YN,
         como_hyperl   = HYPERL_EVER_YN,
         como_cancers  = ifelse(rowSums(.[grep("CANCER_", names(.))]) > 0, 1, 0),
         como_disab    = ifelse(rowSums(.[grep("CATARACT|GLAU|HIP|OSTEO|RA_OA", names(.))]) > 0, 1, 0),
         como_hthl     = ifelse(rowSums(.[grep("HYPERT|HYPERL",  names(.))]) > 0, 1, 0),
         como_resp     = ifelse(rowSums(.[grep("COPD|ASTHMA",  names(.))]) > 0, 1, 0)) %>% 
  mutate_at(vars(starts_with("como_")), factor, labels = c("No", "Yes")) %>% 
  select(starts_with("como_"))

ahs_medic_inc2 <- ahs_medic_inc2 %>% 
  bind_cols(dzdf)

modelvars <- c("bene_age_at_end_2020", 
               "bene_sex_F", 
               "rti_race3", 
               "marital", 
               "educyou", 
               "vegstat", 
               "bmi", 
               "exercise", 
               "sleephrs", 
               "smokecat", 
               "alccat", 
               "como_depress",
               "como_disab", 
               "como_diabetes", 
               # "como_hthl", 
               "como_resp", 
               "como_anemia", 
               "como_kidney", 
               "como_hypoth", 
               "como_cancers",
               "meat_gram",
               "fish_gram",
               "eggs_gram",
               "dairy_gram"
)

# After excluding missing on covariates
# there are 32,045 subjects
complete_cases <- ahs_medic_inc2 %>% 
  select(analysisid, all_of(modelvars)) %>% 
  filter(complete.cases(.)) %>% 
  select(analysisid)

ahs_medic_inc2 <- ahs_medic_inc2 %>% 
  inner_join(complete_cases, by = "analysisid")


# dietary variables -------------------------------------------------------

# Food group variables
# Energy-adjust with zero partition
# By default, variables are log-transformed (excluding zeros)
kcal_adjust <- function(data, var, energy, log=TRUE){
  if (missing(var))
    stop("Need to specify variable for energy-adjustment.")
  if (missing(energy))
    stop("Need to specify energy intake.")
  if (missing(data))
    stop("Need to specify a data frame.")
  df <- eval(substitute(data.frame(y = data$var, ea_y = data$var, kcal = data$energy)))
  count_negative <- sum(df$y < 0, na.rm=TRUE)
  if (count_negative > 0)
    warning("There are negative values in variable.")
  if(log) df$y[df$y > 0 & !is.na(df$y)] <- log(df$y[df$y > 0 & !is.na(df$y)])
  mod <- lm(y ~ kcal, data=df[df$y != 0, ])
  if(log){
    ea <- exp(resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE))
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  else{
    ea <- resid(mod) + mean(df$y[df$y != 0], na.rm=TRUE)
    df$ea_y[!is.na(df$y) & df$y != 0] <- ea
  }
  return(df$ea_y)
}

ahs_medic_inc2$meat_gram_ea      <- kcal_adjust(meat_gram,  kcal, data = ahs_medic_inc2, log = TRUE)
ahs_medic_inc2$fish_gram_ea      <- kcal_adjust(fish_gram,  kcal, data = ahs_medic_inc2, log = TRUE)
ahs_medic_inc2$eggs_gram_ea      <- kcal_adjust(eggs_gram,  kcal, data = ahs_medic_inc2, log = TRUE)
ahs_medic_inc2$dairy_gram_ea     <- kcal_adjust(dairy_gram, kcal, data = ahs_medic_inc2, log = TRUE)
ahs_medic_inc2$nutsseeds_gram_ea <- kcal_adjust(nutsseeds_gram, kcal, data = ahs_medic_inc2, log = TRUE)

# Create quantile groups
# Specify p for other percentile groups
cutQ <- function(x, p=0:4/4, na.rm=FALSE) cut(x, quantile(x, p, na.rm=na.rm), include.lowest=TRUE)

# Quartiles for egg, dairy and nuts/seeds
ea_diet_vars <- c("eggs_gram_ea", "dairy_gram_ea", "nutsseeds_gram_ea")
get_percentile <- function(var, data = ahs_medic_inc2) cutQ(data[[var]], na.rm = TRUE, p = 0:4/4)
test <- lapply(ea_diet_vars, get_percentile)
names(test) <- paste0(ea_diet_vars, "_4")
ahs_medic_inc2 <- cbind(ahs_medic_inc2, test)
rm(test)

ahs_medic_inc2 <- ahs_medic_inc2 %>% 
  mutate(meat_gram_ea_4 = cut(meat_gram_ea, breaks = c(-Inf, 0, 11, 32, Inf), right = TRUE), 
         fish_gram_ea_4 = cut(fish_gram_ea, breaks = c(-Inf, 0,  8.7, 17.4, Inf), right = TRUE))

levels(ahs_medic_inc2$meat_gram_ea_4)  <- c("None", "<11 g/d", "11-<32 g/d", "32+ g/d")
levels(ahs_medic_inc2$fish_gram_ea_4)  <- c("None", "<8.7 g/d", "8.7-<17.4 g/d", "17.4+ g/d")
levels(ahs_medic_inc2$eggs_gram_ea_4)  <- c("<3.6 g/d", "3.6-7.5 g/d", "7.5-<16 g/d", "16+ g/d")
levels(ahs_medic_inc2$dairy_gram_ea_4) <- c("<30 g/d", "30-100 g/d", "100-<236 g/d", "236+ g/d")

```

* Other variables, including demographic, lifestyle, comorbidity, and dietary variables were defined as previously described
  * (Will add these descriptions later)

* Missing values in covariates
  * There were n = 2785 subjects who has at least one missing value on covariates (such as marital status, education, exercise level, sleep hours, smoking and alcohol use, all of them come from AHS questionnaire). These subjects were excluded, resulting in n = 32,045.
  * Among them, there are 11,619 incident cases (36.3%) of CVD

* For those diagnosed with CVD, the mean age at diagnosis was 74.8 years (median 73.6 years)

```{r descriptive_stratified, echo = FALSE}
# Table 1 -----------------------------------------------------------------

# Variables to be included
tablevars <- c("agecat", 
               # "bene_age_at_end_2008", 
               "bene_age_at_end_2020", 
               "bene_sex_F", 
               "rti_race3", 
               "marital", 
               "educyou", 
               "vegstat", 
               "bmicat", 
               "bmi", 
               "exercise", 
               "sleephrs", 
               "smokecat", 
               "alccat", 
               "como_depress",
               "como_disab", 
               "como_diabetes", 
               # "como_hthl", 
               "como_resp", 
               "como_anemia", 
               "como_kidney", 
               "como_hypoth", 
               "como_cancers",
               "meat_gram_ea_4",
               "fish_gram_ea_4",
               "eggs_gram_ea_4",
               "dairy_gram_ea_4"
               )

ahs_medic_inc2 %>% 
  mutate(CVD_YN2 = fct_recode(CVD_YN, "Non-case" = "No", "Case" = "Yes")) %>% 
  CreateTableOne(tablevars, strata = "CVD_YN2", data = ., addOverall = TRUE) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()
```

* Let's look at the same descriptive table only among those with hyperlipidemia. Among n = 32,045, there are 15,522 subjects (48.4%) have diagnosis of hyperlipidemia either before the enrollment or any time during the follow-up.
  * Among these, there are 9,321 incident cases (60.0%) of CVD.

```{r descriptive_stratified_hyperlipidemia, echo = FALSE}
ahs_medic_inc2 %>% 
  filter(HYPERL_YN == "Yes") %>% 
  mutate(CVD_YN2 = fct_recode(CVD_YN, "Non-case" = "No", "Case" = "Yes")) %>% 
  CreateTableOne(tablevars, strata = "CVD_YN2", data = ., addOverall = TRUE) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()
```

## To do

* Let's move on to Cox models: Modeling strategy:
  * Model with dietary pattern, adjusting for demographics and lifestyle variables
  * Model including meat and egg without dietary pattern, adjusting for others
  * Examine meat/egg interaction if necessary
  * Adding further diaetary variables as covariates?
  
* Depending on results, we could change CVD criteria?
  * Get HRs for each of CVD criteria (Ischemic HD, stroke/TIA, aMI, CHF, and afib)
  * For stroke/TIA -- see how vegan's HR changes

* Need to estimate HRs for each level of meat or egg
  * HRs for meat at each level of egg intake
  * HRs for egg at each level of meat intake
  * "Regular" meat or egg intake can be considered as the highest intake group in our sample?
  * "Low" meat or egg intake being the lowest or 2nd lowest group?

